### 2024-06-03 thesis/input_files/analysis.tex
final

\section{Point forecasting}
This section carries out a comparative study between the state of the art methods for point forecasting, introduced in \ref{ch:point}.
As use case, we will consider the task of predicting the electric load from the GEFCom2014 dataset.
In such setting we considered the following regressors
\begin{itemize}
    \item Day
    \item Hour
    \item Month
    \item Day of the week
    \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
    \item Weather temperature
\end{itemize}
Methods will be compared by means of the RMSE, MAE and MAPE scores, see Section \ref{metrics}.
\subsection{Multiple linear regression}
To get started, standard multiple linear regression has been applied, see Figure \ref{fig:mlr_price} for a visualisation. 
\begin{figure}
    \includegraphics[width=\textwidth]{images/mlr_price.png}
    \caption{Multiple linear regression prediction}
    \label{fig:mlr_price}
\end{figure}

@@

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Next, we tried with the autoregressive approach. Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression. This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time. We remind the reader, that the electricity load time series involves both daily and weekly seasonalities. Hence, the need for a more advanced time series model. The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data. Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components. 
Tbats forecast is visualized in Figure \ref{fig:tbats_price}, meanwhile its RMSE on task 1 is 15.08.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/tbats_price.png}
    \caption{Tbats prediction}
    \label{fig:tbats_price}
\end{figure}
From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
\subsection{Prophet}
Following, the prophet model has been considered. 
We get started by considering the base implementation. In such setting, prophet takes as input the time series object and learns its data generating process.
This method achieves on task 1 a RMSE of 23.96, its prediction is visualized in Figure \ref{fig:prophet_price_1}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
    \caption{Prophet prediction}
    \label{fig:prophet_price_1}
\end{figure}
What can be seen is that prophet correctly models the average trend but does not model peaks and lows precisely. 
Next, a more complex model was trained. We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors. Furthermore, we also applied a log transformation to the dependent variables. Doing so, RMSE on task 1 went down to 10.29. Forecast is visualised in Figure \ref{fig:prophet_price2}.
% , moreover figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.png}
    \caption{Prophet model\_2 prediction}
    \label{fig:prophet_price2}
\end{figure}

@@

% quantile regressors
\subsection{Quantile regressors comparison}
Using the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} for Germany and Switzerland, we proceeded to compare between quantile regression algorithms. In this study, we trained our models on the entire 2021 data and then tested them and measured their score on the 2022 data.
Notice, for kernel quantile regression (KQR), the kernel of choice in this setting is the Absolute Laplacian. Additionally notice, the loss function utilized in this section is the normalized pinball loss. 
% normalized by the magnitude of the average load of each country.
Here, we built two models. In the first, the national load is forecasted by taking as predictor the following variables
\begin{itemize}
    \item Weather temperature
    \item Wind speed
\end{itemize}
Results for Germany are reported in table \ref{tab:energy_chart_de} and figure \ref{fig:DE_load_CI} while results for Switzerland can be found in table \ref{tab:energy_chart_ch} and figure \ref{fig:CH_load_CI}.

@@

% kernel comparison
\subsection{Kernel function choice}
This section analyzes which kernel function best fits to the characteristics of the load data.
The kernel functions considered are: Gaussian RBF, Absolute Laplacian, Matern 0.5, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process}{sklearn.gaussian\_process}.
The pool of data for this study has been created by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.
SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modeled by the European Centre for Medium-Range Weather Forecasts (ECMWF).
Therefore, in this experimental study, we used the whole data of $2020$ as the training set and then tested our kernels on the $2021$ data.
Note that there are two kinds of predictions for data from $2021$ onward, one for each of the emission scenarios RCP$4.5$ and RCP$8.5$. In this study, we considered the more conservative RCP$4.5$ scenario data.

The predictors making up the datasets follow:
\begin{itemize}
    \item Direct irradiation: direct normal irradiation
    \item Global radiation: mean global radiation
    \item Hydro reservoir: daily mean power from reservoir plants in MW
    \item Hydro river: daily mean power from run of river plants in MW
    \item Temperature: air temperature
    \item Wind potential: potential wind power production
\end{itemize}
Table \ref{tab:secures_met_de} shows results for Germany, Table \ref{tab:secures_met_ch} is the one for Switzerland and Table \ref{tab:secures_met_at} corresponds to Austria.

@@

\begin{table}[!ht]
    \centering
\caption{Pinball loss kernel comparison for load in Austria (2021)}
\label{tab:secures_met_at}
\begin{adjustbox}{width=\textwidth}
    \begin{tabular}{lllllllllll}
        \toprule
        Quantile & \begin{tabular}[c]{@{}l@{}}Absolute\\ Laplacian\end{tabular} & \begin{tabular}[c]{@{}l@{}}Matern 0.5/\\ Laplacian\end{tabular} & Matern 1.5 & Matern 2.5 & \begin{tabular}[c]{@{}l@{}}Matern $\infty$/\\ Gaussian RBF\end{tabular} & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
        \midrule
        0.1 & \textbf{0.024365} & 0.024592 & 0.024550 & 0.024611 & 0.024831 & 0.026853 & 0.025744 & 0.026004 & 0.027672 & 0.027816 \\
        0.2 & \textbf{0.040971} & 0.041406 & 0.041413 & 0.041487 & 0.041857 & 0.045007 & 0.043247 & 0.043720 & 0.045899 & 0.045879 \\
        0.3 & \textbf{0.052316} & 0.052969 & 0.052888 & 0.052973 & 0.053449 & 0.057880 & 0.055231 & 0.055446 & 0.058572 & 0.058454 \\
        0.4 & \textbf{0.058286} & 0.059018 & 0.058916 & 0.059078 & 0.059592 & 0.065770 & 0.061844 & 0.062495 & 0.066599 & 0.066335 \\
        0.5 & \textbf{0.060344} & 0.061100 & 0.061319 & 0.061541 & 0.062338 & 0.068197 & 0.063864 & 0.071731 & 0.069761 & 0.069574 \\
        0.6 & \textbf{0.058724} & 0.059507 & 0.059674 & 0.059930 & 0.060709 & 0.065828 & 0.062265 & 0.070269 & 0.067179 & 0.066970 \\
        0.7 & \textbf{0.053148} & 0.054189 & 0.054249 & 0.054372 & 0.054916 & 0.058548 & 0.055830 & 0.063015 & 0.059807 & 0.059419 \\
        0.8 & \textbf{0.042740} & 0.043368 & 0.043408 & 0.043510 & 0.043881 & 0.046017 & 0.044384 & 0.049813 & 0.047078 & 0.046662 \\
        0.9 & \textbf{0.026430} & 0.026583 & 0.026781 & 0.026817 & 0.026908 & 0.027050 & 0.026471 & 0.030279 & 0.028031 & 0.027986 \\
        \midrule
        CRPS & \textbf{0.046369} & 0.046970 & 0.047022 & 0.047146 & 0.047609 & 0.051239 & 0.048764 & 0.052530 & 0.052289 & 0.052122 \\
        \bottomrule
    \end{tabular}
\end{adjustbox}
\end{table}
From our tables, it can be concluded that the Absolute Laplacian kernel is the one fitting best to the characteristics of load data.
Additionally, we can conclude that our results are in line with the ones published in \cite{he2017short}. In that study, similarly to what we did in this subsection, the authors carried out a comparitive study concerning the choice of kernels in KQR. Nevertheless notice, that the set of kernel we considered in our study is larger than the one presented in \cite{he2017short}.
Their paper showed evidence of the superiority of the Gaussian RBF among their set of kernel functions. Analogously, we can see from Tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Matern kernels are the best in terms of pinball loss.
The Periodic equipped with period $p=24$ shows good performance. 
After those kernels, we have the Linear and Polynomial kernels.
The worst among all kernels considered are the Cosine and the Sigmoid.
% , which is better suited for binary classification tasks. 
Finally, given that the features dimension is greater than one we have that the Absolute Laplacian and the Matern 0.5 kernels differ. The former kernel is better than the latter and this comes from their distance function. This is because the Manhattan distance is more robust than the Euclidean distance in terms of measuring similarity between objects as data grows in dimensionality \cite{Aggarwal2001}. %(curse of dimensionality)
Those kernel matrices for the Swiss data are compared in Figure \ref{fig:k_matrix_laplacian_comp}.

@@

Our results for this track are reported in Table \ref{tab:pinball loss gefcom2014 price data}. In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}. 
The Gaussian RBF probabilistic prediction for the $4^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualized in Figure \ref{fig:price_task_4_gaussian_rbf}.
Cross validation results are visualised in Figure.

\begin{figure}
    \includegraphics[width=\textwidth]{images/cv_scatter.png}
    \label{fig:cv_scatter}
    \caption{Cross validation price track task 4, Gaussian RBF kernel}
\end{figure}

@@

What can be concluded from Table \ref{tab:pinball loss gefcom2014 price data} is that KQR performs consistently among the top algorithms. Furthermore, It ranks first in one out of the twelve tasks of the price track.
\subsubsection{GEFCom2014 load track} 
The GEFCom2014 load track constitutes a good setting for analysing the performance of KQR in the context of medium term load forecasting (MTLF).
In this track we were faced with the challenge of predicting the load for the next month without the availability of weather temperature forecasts. 
Therefore, the primary task was to first accurately predict the weather temperatures, and then model the load accordingly. Since there were no attributes available for humidity or wind speed, we chose to predict weather temperatures by aggregating historical temperature data across different dimensions such as day, month, and hour. Then we proceeded with building KQR models for the load; we chose the following predictors
\begin{itemize}
    \item Day: the number of the day
    \item Hour
    \item Day of week: an ordinal categorical variable for the day of the week
    \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
    \item w avg: average of weather temperatures across 25 weather stations
\end{itemize}
We built $12$ models, one for each month, with each task model trained on the historical data of the month associated with it.

Table \ref{tab:pinball loss gefcom2014 load data} reports our results for the load track. The top teams for the load forecasting track were Tololo, Adada, Jingrui(Rain) Xie, OxMAth, E.S. Managalova, Ziel Florian, and Bidong Liu; for a breakdown of the attributes of each method, see \cite[Table 6]{hong2016probabilistic}.
Finally, the $90\%$ confidence interval forecast by our model equipped with the Absolute Laplacian kernel for task number $9$, that is the prediction for June, are reported in Figure \ref{fig:load_task_9} and Figuree \ref{fig:boxplot_load_task_9}.
Similarly to the price track, we have that KQR was among the best algorithms also in the load forecasting track.

@@

Among the contributions of this thesis work, we can list
% list contributions:
% - study of kernel methods in EF, both point and probabilistic
% - contribution to less developed area of EF, which is probabilistic forecasting
% - provide open source python package for KQR PyPi, github repo, reproducibility purposes
% - study case for DACH region, along with cleaned dataset for easing collaboration and replication in research
% - extensive comparison among kernel types
% - comparison against state of the art algorithms from GEFCom2014, particularly KQR in medium term load forecasting which has not been deeply analyzed in the literature.
\begin{itemize}
    \item We have studied kernel methods for electricity forecasting both in the point and probabilistic framework.
    \item We provided a Python open source implementation for kernel quantile regression compatible with the sklearn API. The code has been packaged and uploaded to the Python Package Index (PyPI) with the name \href{https://pypi.org/project/kernel-quantile-regression/#2}{kernel-quantile-regression}. The \href{https://github.com/luca-pernigo/ThesisKernelMethods}{github repo} hosting the source code includes also the script implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
    \item We achieved superior performance of kernel quantile regression compared to standard quantile regressor algorithms. 
    \item We created and made available datasets suitable for algorithms benchmarking considering data from the DACH region (Germany, Switzerland and Austria). The format of these data takes inspiration from the popular GEFCom competitions.
    \item Kernel quantile regression extensive comparison between kernel function types.
    \item We compared kernel quantile regression against state of the art probabilistic algorithms in the literature by means of the GEFCom2014 competition.
\end{itemize}

>>>>

\section{Point forecasting}
This section carries out a comparative study between the state of the art methods for point forecasting, introduced in Section \ref{ch:point}.
As use case, we will consider the task of predicting the electric load from the GEFCom2014 dataset.
In such setting we considered the following regressors
\begin{itemize}
    \item Day
    \item Hour
    \item Month
    \item Day of week: an ordinal categorical variable corresponding to the day of the week, e.i.\ Monday=0,$\dots$, Sunday=6
    \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
    \item Weather temperature
\end{itemize}
Methods will be compared by means of the RMSE, MAE and MAPE scores, see Section \ref{metrics}.
\subsection{Multiple linear regression}
To get started, standard multiple linear regression has been applied, see Figure \ref{fig:mlr_price} for a visualisation. 
\begin{figure}
    \includegraphics[width=\textwidth]{images/mlr_price.png}
    \caption{Multiple linear regression prediction}
    \label{fig:mlr_price}
\end{figure}

@@

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Next, we tried with the autoregressive approach. Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression. This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time. We remind the reader, that the electricity load time series involves both daily and weekly seasonalities. Hence, the need for a more advanced time series model. The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data. Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components. 
Tbats forecast is visualised in Figure \ref{fig:tbats_price}, meanwhile its RMSE on task 1 is 15.08.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/tbats_price.png}
    \caption{Tbats prediction}
    \label{fig:tbats_price}
\end{figure}
From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
\subsection{Prophet}
Following, the prophet model has been considered. 
We get started by considering the base implementation. In such setting, prophet takes as input the time series object and learns its data generating process.
This method achieves on task 1 a RMSE of 23.96, its prediction is visualised in Figure \ref{fig:prophet_price_1}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
    \caption{Prophet prediction}
    \label{fig:prophet_price_1}
\end{figure}
What can be seen is that prophet correctly models the average trend but does not model peaks and lows precisely. 
Next, a more complex model was trained. We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors. Furthermore, we also applied a log transformation to the dependent variables. Doing so, RMSE on task 1 went down to 10.29. Forecast is visualised in Figure \ref{fig:prophet_price2}.
% , moreover figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.png}
    \caption{Prophet model\_2 prediction}
    \label{fig:prophet_price2}
\end{figure}

@@

% quantile regressors
\subsection{Quantile regressors comparison}
Using the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} for Germany and Switzerland, we proceeded to compare between quantile regression algorithms. In this study, we trained our models on the entire 2021 data and then tested them and measured their score on the 2022 data.
Notice, for kernel quantile regression (KQR), the kernel of choice in this setting is the Absolute Laplacian. Additionally notice, the loss function utilised in this section is the normalised pinball loss. 
% normalized by the magnitude of the average load of each country.
Here, we built two models. In the first, the national load is forecasted by taking as predictor the following variables
\begin{itemize}
    \item Weather temperature
    \item Wind speed
\end{itemize}
Results for Germany are reported in table \ref{tab:energy_chart_de} and figure \ref{fig:DE_load_CI} while results for Switzerland can be found in table \ref{tab:energy_chart_ch} and figure \ref{fig:CH_load_CI}.

@@

% kernel comparison
\subsection{Kernel function choice}
This section analyses which kernel function best fits to the characteristics of the load data.
The kernel functions considered are: Gaussian RBF, Absolute Laplacian, Matern 0.5, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process}{sklearn.gaussian\_process}.
The pool of data for this study has been created by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.
SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modelled by the European Centre for Medium-Range Weather Forecasts (ECMWF).
Therefore, in this experimental study, we used the whole data of $2020$ as the training set and then tested our kernels on the $2021$ data.
Note that there are two kinds of predictions for data from $2021$ onward, one for each of the emission scenarios RCP$4.5$ and RCP$8.5$. In this study, we considered the more conservative RCP$4.5$ scenario data.

The predictors making up the datasets follow
\begin{itemize}
    \item Direct irradiation: direct normal irradiation
    \item Global radiation: mean global radiation
    \item Hydro reservoir: daily mean power from reservoir plants in MW
    \item Hydro river: daily mean power from run of river plants in MW
    \item Temperature: air temperature
    \item Wind potential: potential wind power production
\end{itemize}
Table \ref{tab:secures_met_de} shows results for Germany, Table \ref{tab:secures_met_ch} is the one for Switzerland and Table \ref{tab:secures_met_at} corresponds to Austria.

@@

\begin{table}[!ht]
    \centering
\caption{Pinball loss kernel comparison for load in Austria (2021)}
\label{tab:secures_met_at}
\begin{adjustbox}{width=\textwidth}
    \begin{tabular}{lllllllllll}
        \toprule
        Quantile & \begin{tabular}[c]{@{}l@{}}Absolute\\ Laplacian\end{tabular} & \begin{tabular}[c]{@{}l@{}}Matern 0.5/\\ Laplacian\end{tabular} & Matern 1.5 & Matern 2.5 & \begin{tabular}[c]{@{}l@{}}Matern $\infty$/\\ Gaussian RBF\end{tabular} & Linear & Periodic & Polynomial & Sigmoid & Cosine \\
        \midrule
        0.1 & \textbf{0.024365} & 0.024592 & 0.024550 & 0.024611 & 0.024831 & 0.026853 & 0.025744 & 0.026004 & 0.027672 & 0.027816 \\
        0.2 & \textbf{0.040971} & 0.041406 & 0.041413 & 0.041487 & 0.041857 & 0.045007 & 0.043247 & 0.043720 & 0.045899 & 0.045879 \\
        0.3 & \textbf{0.052316} & 0.052969 & 0.052888 & 0.052973 & 0.053449 & 0.057880 & 0.055231 & 0.055446 & 0.058572 & 0.058454 \\
        0.4 & \textbf{0.058286} & 0.059018 & 0.058916 & 0.059078 & 0.059592 & 0.065770 & 0.061844 & 0.062495 & 0.066599 & 0.066335 \\
        0.5 & \textbf{0.060344} & 0.061100 & 0.061319 & 0.061541 & 0.062338 & 0.068197 & 0.063864 & 0.071731 & 0.069761 & 0.069574 \\
        0.6 & \textbf{0.058724} & 0.059507 & 0.059674 & 0.059930 & 0.060709 & 0.065828 & 0.062265 & 0.070269 & 0.067179 & 0.066970 \\
        0.7 & \textbf{0.053148} & 0.054189 & 0.054249 & 0.054372 & 0.054916 & 0.058548 & 0.055830 & 0.063015 & 0.059807 & 0.059419 \\
        0.8 & \textbf{0.042740} & 0.043368 & 0.043408 & 0.043510 & 0.043881 & 0.046017 & 0.044384 & 0.049813 & 0.047078 & 0.046662 \\
        0.9 & \textbf{0.026430} & 0.026583 & 0.026781 & 0.026817 & 0.026908 & 0.027050 & 0.026471 & 0.030279 & 0.028031 & 0.027986 \\
        \midrule
        CRPS & \textbf{0.046369} & 0.046970 & 0.047022 & 0.047146 & 0.047609 & 0.051239 & 0.048764 & 0.052530 & 0.052289 & 0.052122 \\
        \bottomrule
    \end{tabular}
\end{adjustbox}
\end{table}
From our tables, it can be concluded that the Absolute Laplacian kernel is the one fitting best to the characteristics of load data.
Additionally, we can conclude that our results are in line with the ones published in \cite{he2017short}. In that study, similarly to what we did in this subsection, the authors carried out a comparative study concerning the choice of kernels in KQR. Nevertheless, notice that the set of kernels we considered in our study is larger than the one presented in \cite{he2017short}.
Their paper showed evidence of the superiority of the Gaussian RBF among their set of kernel functions. Analogously, we can see from Tables \ref{tab:secures_met_de}, \ref{tab:secures_met_ch} and \ref{tab:secures_met_at} that the Matern kernels are the best in terms of pinball loss.
The Periodic equipped with period $p=24$ shows good performance. 
After those kernels, we have the Linear and Polynomial kernels.
The worst among all kernels considered are the Cosine and the Sigmoid.
% , which is better suited for binary classification tasks. 
Finally, given that the features dimension is greater than one we have that the Absolute Laplacian and the Matern 0.5 kernels differ. The former kernel is better than the latter and this comes from their distance function. This is because the Manhattan distance is more robust than the Euclidean distance in terms of measuring similarity between objects as data grows in dimensionality \cite{Aggarwal2001}. %(curse of dimensionality)
Those kernel matrices for the Swiss data are compared in Figure \ref{fig:k_matrix_laplacian_comp}.

@@

Our results for this track are reported in Table \ref{tab:pinball loss gefcom2014 price data}. In this track, the top entries came from the teams: Tololo, Team Poland, GMD, and C3 Green Team; for a breakdown of each method's attributes, see \cite[Table 8]{hong2016probabilistic}. 
The Gaussian RBF probabilistic prediction for the $4^{th}$ July $2013$ zonal price at the $90\%$ confidence interval is visualised in Figure \ref{fig:price_task_4_gaussian_rbf}.
Cross validation results are visualised in Figure \ref{fig:cv_scatter}.

\begin{figure}
    \includegraphics[width=\textwidth]{images/cv_scatter.png}
    \caption{Cross validation price track task 4, Gaussian RBF kernel}
    \label{fig:cv_scatter}
\end{figure}

@@

What can be concluded from Table \ref{tab:pinball loss gefcom2014 price data} is that KQR performs consistently among the top algorithms. Furthermore, it ranks first in one out of the twelve tasks of the price track.
\subsubsection{GEFCom2014 load track} 
The GEFCom2014 load track constitutes a good setting for analysing the performance of KQR in the context of medium term load forecasting (MTLF).
In this track we were faced with the challenge of predicting the load for the next month without the availability of weather temperature forecasts. 
Therefore, the primary task was to first accurately predict the weather temperatures, and then model the load accordingly. Since there were no attributes available for humidity or wind speed, we chose to predict weather temperatures by aggregating historical temperature data across different dimensions such as day, month, and hour. Then we proceeded with building KQR models for the load; we chose the following predictors
\begin{itemize}
    \item Day: the number of the day
    \item Hour
    \item Day of week: an ordinal categorical variable corresponding to the day of the week, e.i.\ Monday=0,$\dots$, Sunday=6
    \item Is holiday: a binary variable for holidays, where holiday=1 and working day=0
    \item w avg: average of weather temperatures across 25 weather stations
\end{itemize}
We built $12$ models, one for each month, with each task model trained on the historical data of the month associated with it.

Table \ref{tab:pinball loss gefcom2014 load data} reports our results for the load track. The top teams for the load forecasting track were Tololo, Adada, Jingrui(Rain) Xie, OxMAth, E.S. Managalova, Ziel Florian, and Bidong Liu; for a breakdown of the attributes of each method, see \cite[Table 6]{hong2016probabilistic}.
Finally, the $90\%$ confidence interval forecast by our model equipped with the Absolute Laplacian kernel for task number $9$, that is the prediction for June, is reported in Figure \ref{fig:load_task_9} and Figure \ref{fig:boxplot_load_task_9}.
Similarly to the price track, we have that KQR was among the best algorithms also in the load forecasting track.

@@

Among the contributions of this thesis work, we can list
% list contributions:
% - study of kernel methods in EF, both point and probabilistic
% - contribution to less developed area of EF, which is probabilistic forecasting
% - provide open source python package for KQR PyPi, github repo, reproducibility purposes
% - study case for DACH region, along with cleaned dataset for easing collaboration and replication in research
% - extensive comparison among kernel types
% - comparison against state of the art algorithms from GEFCom2014, particularly KQR in medium term load forecasting which has not been deeply analyzed in the literature.
\begin{itemize}
    \item We have studied kernel methods for electricity forecasting both in the point and probabilistic framework.
    \item We provided a Python open source implementation for kernel quantile regression compatible with the sklearn API. The code has been packaged and uploaded to the Python Package Index (PyPI) with the name \href{https://pypi.org/project/kernel-quantile-regression/#2}{kernel-quantile-regression}. The \href{https://github.com/luca-pernigo/ThesisKernelMethods}{github repo} hosting the source code includes also the scripts implementing the experiments along with the cleaned datasets; this contribution is intended to forster reproducibility in research.
    \item We achieved superior performance of kernel quantile regression compared to standard quantile regressor algorithms. 
    \item We created and made available datasets suitable for algorithms benchmarking considering data from the DACH region (Germany, Switzerland and Austria). The format of these data takes inspiration from the popular GEFCom competitions.
    \item Kernel quantile regression extensive comparison between kernel function types.
    \item We compared kernel quantile regression against state of the art probabilistic algorithms in the literature by means of the GEFCom2014 competition.
\end{itemize}

### 2024-06-03 thesis/input_files/data_exploratory.tex
final

\subsection{GEFCom2014 load track}
In the load track, contestants were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Any data from the various task folders are shifted by one month between each other.
Also in this track, the first three tasks do not count towards the ranking of the various algorithms.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_\-V2.csv file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
The pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the historical load while Figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally, Figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
These plots motivate the selection of weather temperature as one of the predictors for the load.

@@

% \subsection{GEFCom ETL}
\subsubsection{GEFCom ETL}
The provided data is raw, hence the need for a standardized extract transform and load pipeline. 
Firstly, we drop all Nans and zeros. The provided data for the challenge has a timestamp column not coherent to any time format. Thus, we first pass the data through a method extracting time information: hour, day, month and year. This info is then combined into a datetime object, we selected the ISO format.
Next, since the data does not provide the location for the target utility, the best we can do is selecting the average of weather temperatures as predictor variables. This independent variable is named w avg.
The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming conventions and formatting used in the last track of the challenge. The Python ETL pipeline has since been automated to loop every task folder through Bash scripting.

@@

\subsubsection{EDA}
Following we report the EDA plots alongside with insights that can be extracted from them. We choose to focus the following analysis on the Swiss data only, given that the plots for the German data are pretty similar.
To get started we visualise the historical load data in Figure \ref{fig:CH_historical_load_2021}. As expected, electricity load exhibits a seasonal pattern with highs in the winter and lows during the summer season.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
    \caption{Switzerland historical load (2021)}
    \label{fig:CH_historical_load_2021}
\end{figure}
Next, we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while Figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed. From these plots, we can clearly see there exists a relationship that ties wind speeds and temperatures to load.

@@

Following, we use box plots in order to visualize the relations between categorical independent variables and the dependent one. Such visualisation are reported in Figure \ref{fig:CH_day_of_week_boxplot_2021} and Figure \ref{fig:CH_is_holiday_boxplot_2021}. We can see that workdays involve a higher load and possess a wider range than weekdays and holidays.

>>>>

\subsection{GEFCom2014 load track}
In the load track, contestants were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organisers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Any data from the various task folders are shifted by one month between each other.
Also in this track, the first three tasks do not count towards the ranking of the various algorithms.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_\-V2.csv file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
The pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the historical load while Figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally, Figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
These plots motivate the selection of weather temperature as one of the predictors for the load.

@@

% \subsection{GEFCom ETL}
\subsubsection{GEFCom ETL}
The provided data is raw, hence the need for a standardised extract transform and load pipeline. 
Firstly, we drop all Nans and zeros. The provided data for the challenge has a timestamp column not coherent to any time format. Thus, we first pass the data through a method extracting time information: hour, day, month and year. This info is then combined into a datetime object, we selected the ISO format.
Next, since the data does not provide the location for the target utility, the best we can do is selecting the average of weather temperatures as predictor variables. This independent variable is named w avg.
The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming conventions and formatting used in the last track of the challenge. The Python ETL pipeline has since been automated to loop every task folder through Bash scripting.

@@

\subsubsection{EDA}
Following we report the EDA plots alongside with insights that can be extracted from them. We choose to focus the following analysis on the Swiss data only, given that the plots for the German data are pretty similar.
To get started we visualise the historical load data in Figure \ref{fig:CH_historical_load_2021}. As expected, electricity load exhibits a seasonal pattern with highs in the winter and lows during the summer season.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
    \caption{Switzerland historical load (2021)}
    \label{fig:CH_historical_load_2021}
\end{figure}
Next, we built scatter plots of the independent continous variables against the load as means of visualising correlation between them.
Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while Figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed. From these plots, we can clearly see there exists a relationship that ties wind speeds and temperatures to load.

@@

Following, we use box plots in order to visualise the relations between categorical independent variables and the dependent one. Such visualisations are reported in Figure \ref{fig:CH_day_of_week_boxplot_2021} and Figure \ref{fig:CH_is_holiday_boxplot_2021}. We can see that workdays involve a higher load and possess a wider range than weekdays and holidays.
